Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-14 csf2/main.tex
move around a bunch of defs

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.

\begin{definition}[Robust Compilation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
  %, iff 
  $\isdef$
  \rul{$\forall \pi\in\class, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
  % \end{nscenter}
\end{definition}
% 
Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Thus we write $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.
% \MP{
% 	check that we say classes are hp
% }

@@

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  % \footnotesize
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \ev{Use\ \loc\ n;\sandboxtag;\securitytag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'} \\
                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag;\securitytag'}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag;\securitytag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

  \noindent
  $$
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
                                $$
  % \end{nscenter}
\end{definition}

@@

@@

Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
% \MP{what deos this mean?}
The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
\begin{definition}[Sequential Composition]
  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
\end{definition}

Consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
The following theorem establishes what happens if all these compilation steps were robustly secure with respect to e.g., \gls*{ms}:
The resulting $\irl{Ignition BC}$ code would be \gls*{ms} regardless of the context the binary runs in.

Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

\subsubsection{Securing Optimisations}
Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
% \MPin{survey to cite? this is a bit informal MK: hard to find a survey on different compiler passes... I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
% 
% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
% That way, an order of optimisations can be found by hand that yields the best results.
% \MP{what's the point of the previous 3 sentences?(commented)}
Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.
%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
% 
% \MPin{
% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
%  e.g., remove all deallocations to "optimise" memory management by offloading
%  the work to the OS.

To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.
Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.

For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
They do not demonstrate the scalability of the approach.
Moreover, they are missing the upper and lower compositions.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

>>>>

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{L}$ of the compiled counterpart of $\src{p}$ in such a way that they are ''compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
This is why {\em robust preservation} is a good candidate as a compiler-level security property~\cite{abate2019jour}.

This paper uses a more general notion~\cite{abate2021extacc} that applies to compilers translating from one trace model into another.
To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler. 
\MKin{reference example from case study}
Using this relation, there are two projection functions: (1) the existential image $\tau_\sim\left(\src{\pi}\right)$ and (2) the universal image $\sigma_\sim\left(\trg{\pi}\right)$.
These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties and can be lifted to work with classes, too, by taking the powerset.

For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.

\begin{definition}[Robust Compilation with $\tau_\sim$]\label{def:rtp:tau}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
  %, iff 
  $\isdef$
  \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
  % \end{nscenter}
\end{definition}
% 
Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Because of this, this paper may write $\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\pi}}{\sim}$, even though $\src{\pi}$ is a property and not a class.

Dually to \Cref{def:rtp:tau}, for a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of that class $\trg{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\trg{\pi}$}.

\begin{definition}[Robust Compilation with $\sigma_\sim$]\label{def:rtp:sigma}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\trg{\class}}{\sim}$}
  %, iff 
  $\isdef$
  \rul{$\forall \trg{\pi}\in\trg{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\sigma_\sim\left(\trg{\pi}\right)}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\trg{\pi}}$}.
  % \end{nscenter}
\end{definition}

A degenerate case where $\sim$ coincides with equality, i.e., the trace model is the same for both source and target programs, is the classic definition of robust preservation~\cite{abate2019jour}:

\begin{definition}[Robust Compilation]\label{def:rtp}
  % $\;$\\
  % \vspace{-1em}
  % \begin{nscenter}
  % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
  \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class}$}
  %, iff 
  $\isdef$
  \rul{$\forall \pi\in\class, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\pi}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\pi}$}.
  % \end{nscenter}
\end{definition}

% \MP{
% 	check that we say classes are hp
% }

@@

Notably, real-world compilation chains also perform a series of (sequential) passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in practice, e.g. as in the work of~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion} and many more.
% \MPin{survey to cite? this is a bit informal MK: hard to find a survey on different compiler passes... I can't even find one just for compiler enforced memsafety (that looks good) or for compiler translations in general}.
% 
% Consider the compilation passes of LLVM~\cite{lattner2004llvm}, which are handled in the \texttt{PassManager} class.
% By manipulating the internal state of a \texttt{PassManager} object, the schedule of compiler optimisation passes can be hardcoded.
% That way, an order of optimisations can be found by hand that yields the best results.
% \MP{what's the point of the previous 3 sentences?(commented)}
Consider the following two LLVM optimisation passes: \gls*{cf}, which rewrites constant expressions to the constant they evaluate to, and \gls*{dce}, which removes dead code by rewriting conditional branches.
The order in which \gls*{cf} and \gls*{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
\begin{figure}[!ht]
  \begin{center}
    \begin{tikzpicture}[scale=0.62, every node/.style={transform shape}]
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]
let a = true in
if a then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\large\ttfamily]
if true then
  print "a"
else
  print "b"
\end{lstlisting}};
      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\Large\ttfamily]
print "a"
\end{lstlisting}};% the following is a hack to get my syntax highlighting to work again \end{lstlisting}
      % edges
      \draw[->,very thick] (progunopt) edge[loop left,left] node {\gls*{dce}} (progunopt);
      \draw[->,very thick] (progunopt) edge[sloped,above] node {\gls*{cf}} (progoptcf);
      \draw[->,very thick] (progoptcf) edge[sloped,above] node {\gls*{dce}} (progoptcfdce);
    \end{tikzpicture}
  \end{center}
  \caption{Example program where the level of optimisations differ for one pass of applying \gls*{cf} and \gls*{dce} in any order. %
  Every edge is a compilation pass and the label on the edge states what the pass does, i.e., \gls*{cf} or \gls*{dce}. %
  The source code in the nodes is a glorified compiler intermediate representation and the code gets more optimised towards the right hand side of the figure.}\label{fig:cfdceex}
\end{figure}
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration of the optimisation pipeline~\cite{click1995combining}.

Suppose there are traces $\src{\trace}$ and $\trg{\trace}$ as well as two trace relations $\sim_1$ and $\sim_2$. 
The traces are related $\src{\trace}\sim_1\bullet\sim_2\trg{\trace}$ if there is another trace $\irl{\trace}$ such that $\src{\trace}\sim_1\irl{\trace}$ and $\irl{\trace}\sim_2\trg{\trace}$.

\bul{A cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential image preserves set membership}.
\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tau:tracerel}
  % $$

  \noindent
  \text{\bul{$\wfctau{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\in\src{\class}, \tau_\sim(\src{\Pi})\in\tilde{\tau}_\sim(\src{\class})$}}
  % $$
\end{definition}

\bul{A cross-language relation $\sim$ is well-formed with respect to a class of target-level properties $\trg{\class}$} iff \rul{the universal image preserves set membership}.
\begin{definition}[Well-formedness of Trace Relations for a Class $\trg{\class}$]\label{def:wfc:sig:tracerel}
  % $$

  \noindent
  \text{\bul{$\wfcsig{\sim}{\trg{\class}}$}} \isdef \text{\rul{$\forall \trg{\Pi}\in\trg{\class}, \sigma_\sim(\trg{\Pi})\in\sigma_\sim(\trg{\class})$}}
  % $$
\end{definition}

With these notions of well-formedness, it is now possible to prove the main result that secure compilers in the robust compilation framework~\cite{abate2021extacc} compose {\em sequentially}.

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\tau$]\label{thm:rtpsim:tau}
  If \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wfctau{\sim_1}{\src{\class[_2]}}$}, then \oul{$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

The degenerate case where the cross-language trace relation is equality has a simpler compositionality result:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

%If possible, optimisations are merged to enhance the analysis' result, since even when iterating to a fixpoint the final program may not be as optimised as it could be~\cite{click1995combining}.
Compiler engineers typically try to find an order of optimisations that yields well-optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
\Cref{corr:swappable} justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.
% 
% \MPin{
% 	later on, say that optimisation passes remove elements from traces, so they should be fine with respect to the result with $\tau$.
% 	also, yes, if $\tau$ adds we have a problem, if it removes we don't
% \MKin{ Optimisations that remove elements from traces are not necessarily sound,
%  e.g., remove all deallocations to "optimise" memory management by offloading
%  the work to the OS.
% }
% }

To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations.

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  % \footnotesize
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
                              \ev{Use\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
                              \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

  \noindent
  $$
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag}, \text{ then }m<n
                                \end{array}\right.\right\}
                                $$
  % \end{nscenter}
\end{definition}

@@

\subsubsection{Speculative Safety}

\MKin{fill in the blank based off of spectetor}

@@

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
These works consider languages with different trace models and our technical setup can be adapted to that as long as security properties and their monitors are still defined on the same trace model.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
They do not demonstrate the scalability of the approach.
Moreover, they are missing the upper and lower compositions.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.
